## Known (R)QGIS problems

- `grass7:r.viewshed` does not work under Linux (under Windows it does...)
```{r, eval = FALSE}
data("dem")
data("random_points")

find_algorithms("viewshed")
alg <- "grass7:r.viewshed"
get_usage(alg)
open_help(alg)
# let's find out about the default values
get_args_man(alg)
point <- random_points[sample(1:nrow(random_points), 1), ]
coord <- paste(sf::st_coordinates(point), collapse = ",")
out <- run_qgis(alg, input = dem, coordinates = "796361,8934049",
                output = "out.tif", load_output = TRUE)
# under Linux you might not get an output (I think we should file a bug...)
```

- `grass7:r.watershed`: The algorithm is executed sucessfully but no output is produced though QGIS tells otherwise (issue #81).

```{r, eval = FALSE}
library("RQGIS")
data("dem")
run_qgis("grass7:r.watershed", drainage = "drain.tif", elevation = dem, 
         load_output = TRUE)
```

- `qgis:intersection` produces under Linux the same result when running twice with different input layers. Under Windows it works (issue #74).

```{r, eval = FALSE}
library(sp)
crds <- cbind(x=c(0, 0, 400, 400, 0), y=c(0, 400, 400, 0, 0))
Pl <- Polygon(crds)
ID <- "1"
Pls <- Polygons(list(Pl), ID=ID)
SPls <- SpatialPolygons(list(Pls))
df <- data.frame(value=1, row.names=ID)
SPDF1 <- SpatialPolygonsDataFrame(SPls, df)
crds <- cbind(x=c(0, 0, 400, 400, 0), y=c(0, 400, 400, 0, 0)) - 200
Pl <- Polygon(crds)
ID <- "2"
Pls <- Polygons(list(Pl), ID=ID)
SPls <- SpatialPolygons(list(Pls))
df <- data.frame(value=1, row.names=ID)
SPDF2 <- SpatialPolygonsDataFrame(SPls, df)
crds <- cbind(x=c(0, 0, 400, 400, 0), y=c(0, 400, 400, 0, 0)) + 200
Pl <- Polygon(crds)
ID <- "3"
Pls <- Polygons(list(Pl), ID=ID)
SPls <- SpatialPolygons(list(Pls))
df <- data.frame(value=1, row.names=ID)
SPDF3 <- SpatialPolygonsDataFrame(SPls, df)

library(RQGIS)
# RQGIS::set_env("/Applications/QGIS.app")
set_env(dev = TRUE)
plot(st_geometry(st_as_sf(SPDF1)), xlim = c(-400, 400), ylim = c(-400, 600))
plot(st_geometry(st_as_sf(SPDF2)), add = TRUE)
plot(st_geometry(st_as_sf(SPDF3)), add = TRUE)
test1 <- RQGIS::run_qgis("qgis:intersection", INPUT = SPDF1, INPUT2 = SPDF2, 
                         OUTPUT = file.path(tempdir(), "test1.shp"), 
                         load_output = TRUE)
plot(st_geometry(test1), add = TRUE, col = "red")
test2 <- RQGIS::run_qgis("qgis:intersection", INPUT = SPDF1, INPUT2 = SPDF3, 
                         OUTPUT = file.path(tempdir(), "test2.shp"),
                         load_output = TRUE)
plot(test2, add = TRUE, col = "blue")
identical(test1, test2)
```

- `saga:gridbuffer`: Following example works under Windows with QGIS 2.14 and SAGA 2.1.2 but not with QGIS 2.18 and SAGA 2.3.2

```{r, eval = FALSE}
set_env(dev = FALSE)
data(dem)
run_qgis(alg = "saga:gridbuffer", FEATURES = dem, 
         BUFFERTYPE = "1", 
         DIST = "1")
```