---
title: "RQGIS installation manual and troubleshooting for Windows, Mac, Linux"
author: "Patrick Schratz, Jannes Münchow"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{RQGIS installation instructions and troubleshooting for Windows, Mac, Linux"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Vignette Info

This Vignette shows how to set up your local QGIS, SAGA and GRASS environments 
on your local PC to get RQGIS running smoothly. We present example combinations 
of different software versions for each operating system (Windows, Mac, Linux) 
that will interact nicely with each other and RQGIS. Other combinations of 
software version combination of QGIS, SAGA and GRASS than the ones presented 
here might also work without problems. However, for time reasons we can not test
all possible combinations and if you encounter problems it is recommended to use
the newest stable version of all softwares.  



## Windows

### QGIS (OSGEO4W)

For Windows, a convenient installer is available, provided by the 
['Open Source Geospatial Foundation'](http://www.osgeo.org/). This installer
installs QGIS, GRASS, SAGA and all required libraries for you in a way that 
enables you to run QGIS and RQGIS with full power, meaning with all algorithms
QGIS provided. It also takes care of all required bindings. Please download the 
specific installer for your system architecture 
[under this link](http://trac.osgeo.org/osgeo4w/).  

**Option 1** (# working with 840 algs)  

1. Select "Express Installation" 
2. Check "QGIS, GRASS and GDAL"
3. Run!

<span style="color:red">**Option 2:** (not working right now - various erros pointing to missing .dll´s)

1. Select "Advanced" Installation 
2. Check "Download from the Internet"  
3. Optionally select "All Users" oder "Just me"  
4. Leave install directory default  
5. Search for "GDAL" and install all entries 
6. Search for "QGIS"
6.1 Under "Desktop" install "qgis:QGIS Desktop"  
6.2 Under "Libs" install all entries  
7. Search for "GRASS"  
7.1 Under "Desktop" install "grass: GRASS GIS - stable release"  
7.2 Under "Libs" install all entries  
8. Search for "SAGA" and install it  
9. Search for "msys" and install it  
10. Search for "qt4" and install all entries having "qt4" in their name  
11. Search for "python"   
11.1 Under "Commandline_Utilities" install "python-core" and "python-tools"  
11.2 Under "Libs" makes sure that "gdal-dev-python" is installed  

Click on "Continue" and install all selected items. Be patient, it may take 
some time depending on the number of new installation/updates. </span>


#### Troubleshooting
If you encounter the following error when trying to execute `find_algorithms()`:  

**Error #1:**   

``` {r, eval = FALSE}
find_algorithms()
Trying to find OSGeo4W on your C: drive.
 Error in set_env() : There are several QGIS installations on your system:
C:\OSGeo4W64\bin\qgis.bat
C:\Program Files\QGIS Essen\bin\qgis.bat 
```
*Solution: Make sure to delete the QGIS installation which is !NOT! installed* 
*under "C:\\OSGEO4W\\"*

----

**Error #2:**  (advanced installation)

```{r, eval = FALSE}
find_algorithms()
Trying to find OSGeo4W on your C: drive.
[1] "Traceback (most recent call last):"                                                        
[2] "  File \"py_cmd.py\", line 3, in <module>"                                                 
[3] "    from qgis.core import *"                                                               
[4] "  File \"C:\\OSGEO4~1\\apps\\qgis\\python\\qgis\\__init__.py\", line 36, in <module>"      
[5] "    from qgis.core import QgsFeature, QgsGeometry"                                         
[6] "  File \"C:\\OSGEO4~1\\apps\\qgis\\python\\qgis\\core\\__init__.py\", line 34, in <module>"
[7] "    from qgis._core import *"                                                              
[8] "ImportError: DLL load failed: Das angegebene Modul wurde nicht gefunden."                  
attr(,"status")
[1] 1
Warning message:
running command 'batch_cmd.cmd' had status 1 
```

*Solution: This error appears when there´s something wrong with your local QGIS installation.*
*It is recommended to uninstall all QGIS installations on your PC and install* 
*QGIS using 'Option #1'.*  


## Mac
Number of algorithms: 

- 875 (QGIS + SAGA installation)

### QGIS
Please follow the instructions provided 
[here](https://www.qgis.org/de/site/forusers/download.html). We recommend to 
install the latest 'current' version. 

### SAGA
First, download Fink from Fink Project and [install it](http://www.finkproject.org/doc/bundled/install-first.php).
Then download the following .info file for SAGA:   [http://fink.cvs.sourceforge.net/viewvc/fink/experimental/babayoshihiko/finkinfo/saga.info](http://fink.cvs.sourceforge.net/viewvc/fink/experimental/babayoshihiko/finkinfo/saga.info)  
Copy this file in /sw/fink/dists/local/main/finkinfo. You will need root access. Finally, run
fink install saga
This will automatically download and install necessary libraries as well.

Source: [https://sourceforge.net/p/saga-gis/wiki/Compiling%20SAGA%20on%20Mac%20OS%20X/](https://sourceforge.net/p/saga-gis/wiki/Compiling%20SAGA%20on%20Mac%20OS%20X/)

#### Troubleshooting
Warnings appears when running any RQGIS command. Does not affect usage.  
<span style="color:red">!Unresolved! (!appears also under Linux after OTB configuration!)
```{r, eval = FALSE}
QSqlDatabase: QSQLITE driver not loaded
QSqlDatabase: available drivers: 
QSqlQuery::prepare: database not open
```

## Linux
Number of algorithms: 

- 781 (QGIS + SAGA installation)
- 882 (+ Orfeo Toolbox)


### QGIS
For Debian/Ubuntu, please follow the installation instructions found under this
[link](https://www.qgis.org/de/site/forusers/alldownloads.html). Note that if 
you install QGIS from the built-in Software managers/repositories, you will most 
likely get a rather old QGIS version as these repositories only getting updated 
quite sparsely (especially if your running older Ubuntu builds like 14.04). 
We did not test other Linux derivates at the moment.

You can also just use this simplified guide:

1. Open a terminal window
2. Type `sudo gedit /etc/apt/sources.list`
3. Move to the end of the document and add the following lines:   
  `#QGIS`  
  `deb https://www.qgis.org/debian xenial main` "xenial" here refers to 
  Ubuntu build name which is "Xenial Xerus" for Version 16.04. For Ubuntu 14.04
  replace "xenial" with "trusty". Save it and exit. 
4. In the terminal window: 
add `wget -O - http://qgis.org/downloads/qgis-2015.gpg.key | gpg --import` 
followed by `gpg --export --armor 3FF5FFCAD71472C4 | sudo apt-key add -`
5. Terminal: `sudo apt-get update` 
6. Terminal: `sudo apt-get install qgis python-qgis qgis-plugin-grass`

You basically added the most recent stable QGIS repository to your list of
repositories. Everytime you are running `sudo apt-get update` now, Ubuntu checks for
new QGIS updates. 

### RQGIS
**Compiling RQGIS on a fresh Linux installation:**

After installing R, you can install RQGIS using the `devtools` package. This packages
depends on several libraries which you should install a priori.
We recommend to use the "Synaptic" package manager for this. To install synaptic, please 
use the built-in "Ubuntu Software" app or any other built-in program of your Linux installation.
Search for "libssl-dev","libcurl4-openssl-dev" "libssh2-1-dev" and install all dependencies. Afterwards install the R-package "roxygen2" and then "devtools". If that succeeded, you can install RQGIS
using `devtools::install_github("jannes-m/RQGIS")` and load it with `library("RQGIS")`.

### Enabling additional providers and their algorithms

#### SAGA
To use SAGA functions within (R)QGIS, you naturally need to install SAGA GIS. To install the most recent SAGA version, simply execute the following lines in a terminal window: 

```{bash, eval = FALSE}
sudo add-apt-repository ppa:johanvdw/saga-gis  
sudo apt-get update
sudo apt-get install saga
```

#### Orfeo Toolbox
Orfeo Toolbox has to be installed on your own on Linux systems. To do so, visit [this site](https://www.orfeo-toolbox.org/download/) and save the .run script on your Desktop. 
OTB needs to be installed in the first place in a folder where you have access to. Later on we will the important folders to their desired location. 

We recommend to create a folder called "Applications" under `/home/<YourUserName>/`. To do so, simply execute `mkdir /home/<YourUserName/Applications/`. Then move your .run script to this location by `mv /home/<YourUserName>/Desktop/<filename> /home/<YourUserName>/Applications`. Now run the script from the terminal using `sh /usr/<NameOfScript>`. To ensure that (R)QGIS can find the algorithms we have to move the contents of your installation to the specfic directories. Afterwards you can delete the first installation under [...]/Applications.

(R)QGIS assumes the "OTB applications folder" to be located at "/usr/lib/otb/applications". Call `sudo mv /home/<YourUserName>/Applications/<YourOTBInstallation>/lib/otb /usr/lib/`. This command will move our applications folder (located under [...]/otb) to the desired location. We need "sudo" rights to move a folder there. 

Last, we need to move the "OTB command line tools" contents to their desired place. Call `cp -a /home/<YourUserName>/Applications/<YourOTBInstallation>/bin/. /usr/bin`. 

Now QGIS should list the OTB tools under "Processing Toolbox" and RQGIS should list them when calling `find_algorithms()`. 

For further reading please see the official [QGIS manual](/home/<YourUserName>/Applications/<YourOTBInstallation>/) for OTB installation under Linux. 