---
title: "RQGIS install instructions for Windows, Mac, Linux"
author: "Jannes Münchow, Patrick Schratz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"Install instructions for Windows/Mac/Linux"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Vignette Info

This Vignette shows how to set up your local QGIS, SAGA and GRASS environments 
on your local PC to get RQGIS running smoothly. We present example combinations 
of different software versions for each operating system (Windows, Mac, Linux) 
that will interact nicely with each other and RQGIS. Other combinations of 
software version combination of QGIS, SAGA and GRASS than the ones presented 
here might also work without problems. However, for time reasons we can not test
all possible combinations and if you encounter problems it is recommended to use
the newest stable version of all softwares.  



## Windows

## Mac


## Linux

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
